---
title: "Effects of Skagit Precipitation Patterns on Steelhead Smolt Outmigration"
output: ioslides_presentation
---

## Required Packages

```{r, message=FALSE}
library(tidyverse)
```

Specific package use will include

* Dplyr
* GGPlot

## Data 1: Monthly Discharge

```{r, echo=FALSE}
avg_monthly_discharge <- read.csv("../clean_data/avg_monthly_discharge.txt")
avg_monthly_discharge <- avg_monthly_discharge %>% rename("1" = X1, "2" = X2, "3" = X3, "4" = X4, "5" = X5, "6" = X6, "7" = X7, "8" = X8, "9" = X9, "10" = X10, "11" = X11, "12" = X12)
knitr::kable(avg_monthly_discharge[1:5, 1:7])
``` 

Holds the average discharge for each month of the year from 1941 to 2018.

* Each variable header represents the month of the year
* Discharge is in cubic feet per second

## Data 2: Daily Discharge

```{r, echo = FALSE}
avg_daily_discharge <- read.csv("../data/avg_daily_discharge.csv")
knitr::kable(avg_daily_discharge[1:5, 1:4])
```

Each mean value represents the average discharge for that day of the year.

* Discharge given in cubic feet per second
* Mean calculated from data dating 1941 to 2018

## Data 3: Smolt Trap Catch

```{r, echo=FALSE}
yearly_smolt <- read.csv("../data/steelhead_smolt_trapped.csv")
knitr::kable(yearly_smolt[1:5, 1:3])
```

Each smolt value represents the number of smolt caught in a specific type of trap for that year. Trap catches steelhead smolt outmigrants on Skagit river with data from 1990-2007.

## Annual Skagit Hydrograph

```{r, echo=FALSE}
ggplot(avg_daily_discharge)+
  geom_point(mapping = aes(x=day.of.year, y=mean))+
  labs(x="day of year", y= "average discharge (cfs)", title = "Average Annual Hydrograph")
```

Plot graphs average daily discharge values against day of year to show average year of discharge for Skagit river. 

## Annual Skagit Hydrograph cont.

Plot indicates a precipitation pattern dominated by both rain and snow.

* Peak in fall result of high rains

* Peak in winter result of warmer weather melting accumulated snow pack in higher elevations
* Snow melt brings not just high discharge but also cool water temperatures

## Image

![image_name](location)