---
title: "01_data_import"
output: html_document
---
# Required Packages

```{r, message=FALSE}
library(tidyverse)
```


# Import Data


### Monthly Average Discharge

```{r}
avg_monthly_discharge <- read.csv("../data/avg_monthly_discharge.csv")
avg_monthly_discharge <- mutate(avg_monthly_discharge, year = X)
head(avg_monthly_discharge)
```

For more clear organization, a year column was created for easier location. For every other header, the number following the "X" represents the month of the year. For example, "X1" is the column for January.

### Daily Average Discharge

```{r}
avg_daily_discharge <- read.csv("../data/avg_daily_discharge.csv")
head(avg_daily_discharge)
```


### Steelhead Smolt Trapped per Year

```{r}
yearly_smolt <- read.csv("../data/steelhead_smolt_trapped.csv")
head(yearly_smolt)
```


# Average Hydrograph

```{r}
ggplot(avg_daily_discharge)+
  geom_point(mapping = aes(x=day.of.year, y=mean))+
  labs(x="day of year", y= "average discharge (cfs)", title = "Average Annual Hydrograph")
```

The plot uses the average discharge for each day with data from 1940 to 2018. Each point represents the average discharge in cubic feet per second for that specific day of the year.