---
title: "02_initial_plots"
output: html_document
---

```{r}
library(tidyverse)
```

# Data Review

All the data used for analyis:

### Average Monthly Discharge

```{r}
avg_monthly_discharge <- read.csv("../clean_data/avg_monthly_discharge.txt")
avg_monthly_discharge <- avg_monthly_discharge %>% rename("1" = X1, "2" = X2, "3" = X3, "4" = X4, "5" = X5, "6" = X6, "7" = X7, "8" = X8, "9" = X9, "10" = X10, "11" = X11, "12" = X12)
head(avg_monthly_discharge)
```

### Daily Average Discharge

```{r}
avg_daily_discharge <- read.csv("../data/avg_daily_discharge.csv")
head(avg_daily_discharge)
```

### Steelhead Smolt Trapped per Year

```{r}
yearly_smolt <- read.csv("../data/steelhead_smolt_trapped.csv")
head(yearly_smolt)
```

# Average Hydrograph


```{r}
ggplot(avg_daily_discharge)+
  geom_point(mapping = aes(x=day.of.year, y=mean))+
  labs(x="day of year", y= "average discharge (cfs)", title = "Average Annual Hydrograph")
```

The plot uses the average discharge for each day with data from 1940 to 2018. Each point represents the average discharge in cubic feet per second for that specific day of the year.

Peaks around May and November/December indicate watershed dominated by both rain and snow. Spring peak around May and into June originates from snow melt at higher elevations as temperatures rise with the season.

# Late Spring Discharge

```{r, echo = FALSE}
late_spring_discharge <- subset(avg_monthly_discharge, select = c(6, 7,year)) %>%
  rename(may = "5", june = "6")
late_spring_discharge <- mutate(late_spring_discharge, late_spring = may + june)
knitr::kable(late_spring_discharge[1:5, 1:4])
```

This creates a new table with just the average monthly discharges for may and june as well as a sum for the two months. May and June are the peak of the snow melt discharge of spring so this table provides the data to analyze that peak.

# Late Spring Discharge vs Smolt Outmigration

```{r}
discharge_and_smolt <- inner_join(late_spring_discharge, 
                                  yearly_smolt, by = "year")
```
```{r, echo=FALSE}
knitr::kable(discharge_and_smolt[1:4,1:6])
```

The code here joins the late spring discharge table and the smolt outmigration table for analysis.

```{r, echo=FALSE}
ggplot(discharge_and_smolt)+
  geom_point(mapping = aes(x=late_spring, y=smolt..scoop.))+
  labs(x="discharge (cfs)", y="smolt catch #", title = "Discharge vs Smolt Catch")+
  geom_smooth(mapping = aes(x=late_spring, y=smolt..scoop.), method = "lm")
```

This code plots smolt outmigration as a function of late spring discharge. It also ncludes trendline to indicate a potential relationship between smolt outmigration and higher spring discharges.

```{r,echo=FALSE}
smolt_regression <- lm(smolt..scoop. ~ late_spring, 
                       data = discharge_and_smolt)
summary(smolt_regression)
```

A linear regression model here finds no relationship between smolt outmigration and late spring discharge.

# Potential Climate Change Analysis

In the interest of future research, this study also looked at the changes in late spring discharge over time and the potential effects of climate change.

```{r, echo=FALSE}
ggplot(late_spring_discharge)+
  geom_point(mapping = aes(x=year, y=late_spring))+
  geom_smooth(mapping = aes(x=year, y=late_spring), method="lm")+
  labs(y="late spring discharge (cfs)", title="Late Spring Discharge over Time")
```
```{r,echo=FALSE}
late_spring_discharge_regression <- lm(late_spring ~ year, 
                                       data = late_spring_discharge)
summary(late_spring_discharge_regression)
```

The plot and linear regression model indicate a decrease in late spring discharge over time, implying climate change has reduced the snow pack in the Skagit basin that creates the spring peak flows.

# Future Research

More data on a shorter timescale

* More years of smolt catch (2007-present)
* Weekly/daily catch numbers

Threat of climate change

* important to understand relationship
* Timing/intensity of discharge likely to change

# Challenges

* Importing data without commas
* Resizing images to fit in the presentation
* Completing and displaying results of linear regression
* Pulling correct columns for analysis tables



